---
title: "Augusta University: STAT 7630"
subtitle: "Applied Linear Models"
author: "Dustin Pluta"
date: "2024 JAN 09"
output:
  xaringan::moon_reader:
    css: [metropolis-fonts, "my-theme.css"]
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: haddock
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.align="center", cache = TRUE, warning = FALSE)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
```

# Lecture 4

- Confidence Intervals

- Multiple Testing

---

# Confidence Intervals

- __Definition__ A $(1 - \alpha)100\%$ __confidence interval__ for scalar parameter $\theta$ 
constructed from a sample $X$ is an interval $(L(X), U(X)) \subset (-\infty, \infty)$ such that the probability of such an interval containing $\theta$ is $(1 - \alpha)100\%$.

- __Definition__ We refer to $(1 - \alpha)100\%$ as the __coverage__ or __coverage probability__ of the confidence interval.

- The _coverage_ of the interval is the expected proportion of times that the CI will contain the true value $\theta$.  This means that the $(1 - \alpha)100\%$ confidence level is a probability statement with respect to the _distribution of confidence intervals_ constructed this way.

---

# Confidence Intervals

- __Note__ In the frequentist framework, we assume paramters are _fixed_, not random.  

- Consequently, for a given sample, the corresponding CI either contains $\theta$, or it does not.  

- In other words, once the sample is drawn, there is no more randomness, 
and we cannot make probability statements about the specific CI we constructed.


---

# Confidence Intervals

### Example: One-sample Normal, unknown mean and variance

Consider an iid sample $X_i \sim N(\mu, \sigma^2)$.  We wish to construct a 95\% confidence interval for $\mu$.

For the one-sample t-test, we use test statistic

$$T = \frac{\bar X - \mu_0}{s / \sqrt{n}} \overset{H_0}\sim t(n - 1).$$


---

# Confidence Intervals

### Example: One-sample Normal, unknown mean and variance

To form a 95% percent confidence interval for $\mu$, we invert this hypothesis test.  

Note that if we replace $\mu_0$ with the true parameter value $\mu$, $T \sim t(n - 1)$.  Let $t_{1 - \alpha / 2}(n - 1)$ be the $(1 - \alpha / 2)$ percentile of $t$.

$$\begin{aligned}
&P(-t_{1 - \alpha / 2}(n - 1) < T < t_{1 - \alpha / 2}) = 1 - \alpha\\
&P(-t_{1 - \alpha / 2}(n - 1) < \frac{\bar X - \mu}{s / \sqrt{n}} < t_{1 - \alpha / 2}) = 1 - \alpha\\
&P\left(\bar X - \frac{s}{\sqrt{n}} t_{1 - \alpha / 2}(n - 1) < \mu < \bar X + \frac{s}{\sqrt{n}} t_{1 - \alpha / 2}(n - 1)\right) = 1 - \alpha\\
\end{aligned}$$

---

# Confidence Intervals

### Example: One-sample Normal, unknown mean and variance

```{r, echo=TRUE}
set.seed(123)
n <- 20
mu <- 2
sigma <- 1
x <- rnorm(n, mu, sigma)
x_bar <- mean(x)
s <- sd(x)
```

---

# Confidence Intervals

### Example: One-sample Normal, unknown mean and variance


```{r, echo=FALSE}
hist(x)
abline(v = x_bar, col = "red")
```


---

# Confidence Intervals

### Example: One-sample Normal

```{r, echo=TRUE}
alpha <- 0.05
lwr <- x_bar - s / sqrt(n) * pt(1 - alpha / 2, n - 1)
upr <- x_bar + s / sqrt(n) * pt(1 - alpha / 2, n - 1)
```

```{r}
cat(lwr, upr)
```

We see this confidence interval contains the true value $\mu = 2$.

- What happens if we sample $X$ many times and form a confidence interval for each?

---

# Confidence Intervals

### Example: One-sample Normal

```{r, echo=TRUE}
conf_int_simulation <- function(n_sims, n, mu, sigma, alpha = 0.05) {
  results <- data.frame(lwr = rep(NA, n_sims), upr = rep(NA, n_sims))
  
  for (k in 1:n_sims) {
    x <- rnorm(n, mu, sigma)
    x_bar <- mean(x)
    s <- sd(x)
    lwr <- x_bar - s / sqrt(n) * qt(1 - alpha / 2, n - 1)
    upr <- x_bar + s / sqrt(n) * qt(1 - alpha / 2, n - 1)
    results[k, ] <- c(lwr, upr)
  }
  results$sim_index <- 1:nrow(results)
  results$covers <- (mu > results$lwr) & (mu < results$upr)

  return(results)
}
```

---

# Confidence Intervals

```{r, echo=TRUE}
set.seed(123)
n_sims <- 100
n <- 30
mu <- 2
sigma <- 1
alpha <- 0.05

sim_results <- conf_int_simulation(n_sims, n, mu, sigma, alpha)
head(sim_results)
```


```{r, echo=FALSE}
cat("Coverage: ", mean(sim_results$covers))
```

---

# Confidence Intervals


```{r, echo = FALSE}
ggplot(sim_results) + 
  geom_segment(aes(y = sim_index, yend = sim_index, x = lwr, xend = upr, color = covers), size = 1.25) + 
  geom_segment(y = 0, yend = n_sims, x = mu, xend = mu, size = 1.25, linetype = 2)
```

---

# Confidence Intervals

### Example: One-sample Normal

```{r, echo=TRUE}
set.seed(123)
n_sims <- 10000
n <- 30
mu <- 2
sigma <- 1
alpha <- 0.05

sim_results <- conf_int_simulation(n_sims, n, mu, sigma, alpha)
mean(sim_results$covers)
```

---

# Confidence Intervals

### Example: One-sample Normal, unknown mean and variance


```{r, echo=FALSE}
plot(cumsum(sim_results$covers) / 1:n_sims, ty = "l", xlab = "Coverage", ylab = "# of Simulations")
abline(h = 1 - alpha)
```

---

# Multiple Testing


```{r}
set.seed(123)
n_sims <- 40
n <- 25
mu1 <- 0
mu2 <- 0
sigma <- 1
alpha <- 0.05
```

```{r, echo=TRUE}
sim_results1 <- conf_int_simulation(n_sims, n, mu1, sigma, alpha)
sim_results2 <- conf_int_simulation(n_sims, n, mu2, sigma, alpha)
combined_results <- data.frame(covers1 = sim_results1$covers, covers2 = sim_results2$covers, sim = 1:n_sims)
combined_results$covers_all <- combined_results$covers1 & combined_results$covers2
```

```{r, echo = FALSE}
cat("Combined coverage probability: ", mean(combined_results$covers_all), "\n")
cat("Group 1 coverage: ", mean(sim_results1$covers), "\n")
cat("Group 2 coverage: ", mean(sim_results2$covers), "\n")
```

---

# Multiple Testing

```{r, echo=FALSE}
sim_results <- bind_rows(sim_results1, sim_results2)
sim_results$grp <- rep(c(1, 2), each = n_sims)
sim_results$sim <- 1:n_sims
```


```{r, echo=FALSE}
ggplot(sim_results) + 
  geom_segment(aes(y = sim, yend = sim, x = lwr, xend = upr, color = covers), size = 1.3) + 
  xlab("CI for mu") + 
  facet_grid(cols = vars(grp))
```


---

# Multiple Testing

```{r}
set.seed(123)
n_sims <- 40
n <- 30
mu1 <- 0
mu2 <- 0
mu3 <- 0
sigma <- 1
alpha <- 0.05
```

```{r, echo=FALSE}
sim_results1 <- conf_int_simulation(n_sims, n, mu1, sigma, alpha)
sim_results2 <- conf_int_simulation(n_sims, n, mu2, sigma, alpha)
sim_results3 <- conf_int_simulation(n_sims, n, mu3, sigma, alpha)
combined_results <- data.frame(covers1 = sim_results1$covers, covers2 = sim_results2$covers, covers3 = sim_results3$covers)
combined_results$covers_all <- combined_results$covers1 & combined_results$covers2 & combined_results$covers3
```

```{r, echo = FALSE}
cat("Combined coverage probability: ", mean(combined_results$covers_all), "\n")
cat("Group 1 coverage: ", mean(sim_results1$covers), "\n")
cat("Group 2 coverage: ", mean(sim_results2$covers), "\n")
cat("Group 3 coverage: ", mean(sim_results3$covers))
```

---

# Multiple Testing

```{r}
set.seed(123)
n_sims <- 10000
n <- 30
mu1 <- 0
mu2 <- 0
mu3 <- 0
mu4 <- 0
sigma <- 1
alpha <- 0.05
```

```{r, echo=FALSE}
sim_results1 <- conf_int_simulation(n_sims, n, mu1, sigma, alpha)
sim_results2 <- conf_int_simulation(n_sims, n, mu2, sigma, alpha)
sim_results3 <- conf_int_simulation(n_sims, n, mu3, sigma, alpha)
sim_results4 <- conf_int_simulation(n_sims, n, mu4, sigma, alpha)
combined_results <- data.frame(covers1 = sim_results1$covers, covers2 = sim_results2$covers, covers3 = sim_results3$covers, covers4 = sim_results4$covers)
combined_results$covers_all <- combined_results$covers1 & combined_results$covers2 & combined_results$covers3 & combined_results$covers4
```

```{r, echo = FALSE}
cat("Combined coverage probability: ", mean(combined_results$covers_all), "\n")
cat("Group 1 coverage: ", mean(sim_results1$covers), "\n")
cat("Group 2 coverage: ", mean(sim_results2$covers), "\n")
cat("Group 3 coverage: ", mean(sim_results3$covers), "\n")
cat("Group 4 coverage: ", mean(sim_results4$covers))
```

---

# Multiple Testing

## Bonferroni Correction

- When conducting $K$ hypothesis tests simultaneously, the _Bonferroni method_ adjusts the nominal significance level 
from $\alpha$ to $\alpha / K$ on each individual test, in order to maintain the correct Type I error globally, referred to 
as the *familywise error rate* (FWER).  

- When constructing multiple confidence intervals for the one-sample normal case, this has the form

$$\bar X \pm \frac{s}{\sqrt{n}}t_{1 - \alpha / (2K)}(n - 1).$$

- The name of this method is from the Bonferroni inequality from probability theory, which can be stated as

$$P\left(\cup_{k = 1}^K A_k\right) \leq \sum_{k = 1}^K P(A_k),$$

for some set of events $A_k, k = 1, \dots, K$.

---

# Multiple Testing

## Bonferroni Correction

```{r, echo=TRUE}
bonferroni_simulation <- function(n_sims, n, mu, sigma, K, alpha = 0.05) {
  results <- data.frame(lwr = rep(NA, n_sims), upr = rep(NA, n_sims))
  
  for (k in 1:n_sims) {
    x <- rnorm(n, mu, sigma)
    x_bar <- mean(x)
    s <- sd(x)
    lwr <- x_bar - s / sqrt(n) * qt(1 - alpha / (2 * K), n - 1)
    upr <- x_bar + s / sqrt(n) * qt(1 - alpha / (2 * K), n - 1)
    results[k, ] <- c(lwr, upr)
  }
  results$sim_index <- 1:nrow(results)
  results$covers <- (mu > results$lwr) & (mu < results$upr)

  return(results)
}
```

---

# Multiple Testing

## Bonferroni Correction

```{r}
set.seed(12)
n_sims <- 40
n <- 25
mu1 <- 0
mu2 <- 0
sigma <- 1
K <- 2
alpha <- 0.05
```

```{r, echo=FALSE}
sim_results1 <- bonferroni_simulation(n_sims, n, mu1, sigma, K, alpha)
sim_results2 <- bonferroni_simulation(n_sims, n, mu2, sigma, K, alpha)
combined_results <- data.frame(covers1 = sim_results1$covers, covers2 = sim_results2$covers, sim = 1:n_sims)
combined_results$covers_all <- combined_results$covers1 & combined_results$covers2
```

```{r, echo = FALSE}
cat("Combined coverage probability: ", mean(combined_results$covers_all), "\n")
cat("Group 1 coverage: ", mean(sim_results1$covers), "\n")
cat("Group 2 coverage: ", mean(sim_results2$covers), "\n")
```

---

# Multiple Testing

## Bonferroni Correction

```{r, echo=FALSE}
sim_results <- bind_rows(sim_results1, sim_results2)
sim_results$grp <- rep(c(1, 2), each = n_sims)
sim_results$sim <- 1:n_sims
```


```{r, echo=FALSE}
ggplot(sim_results) + 
  geom_segment(aes(y = sim, yend = sim, x = lwr, xend = upr, color = covers), size = 1.3) + 
  xlab("CI for mu") + 
  facet_grid(cols = vars(grp))
```

---

# Multiple Testing

## Bonferroni Correction

```{r}
set.seed(12)
n_sims <- 40
n <- 25
mu1 <- 0
mu2 <- 0
sigma <- 1
K <- 2
alpha <- 0.05
```

```{r, echo=TRUE}
set.seed(12)
bonf_sim_results1 <- bonferroni_simulation(n_sims, n, mu1, sigma, K, alpha)
bonf_sim_results2 <- bonferroni_simulation(n_sims, n, mu2, sigma, K, alpha)
set.seed(12)
sim_results1 <- conf_int_simulation(n_sims, n, mu1, sigma, alpha)
sim_results2 <- conf_int_simulation(n_sims, n, mu2, sigma, alpha)
```


---

# Multiple Testing

## Bonferroni Correction

```{r, echo=FALSE}
sim_results <- bind_rows(sim_results1, sim_results2)
sim_results$grp <- rep(c(1, 2), each = n_sims)
sim_results$sim <- 1:n_sims
sim_results$lwr_bonf <- c(bonf_sim_results1$lwr, bonf_sim_results2$lwr)
sim_results$upr_bonf <- c(bonf_sim_results1$upr, bonf_sim_results2$upr)
```


```{r, echo=FALSE}
ggplot(sim_results) + 
  geom_segment(aes(y = sim, yend = sim, x = lwr_bonf, xend = upr_bonf), color = "darkgreen", alpha = 0.75, size = 1.3) + 
  geom_segment(aes(y = sim, yend = sim, x = lwr, xend = upr, color = covers), size = 1.3) + 
  xlab("CI for mu") + 
  facet_grid(cols = vars(grp))
```


---

# Multiple Testing

## Bonferroni Correction


```{r}
set.seed(123)
n_sims <- 5000
n <- 30
mu1 <- 0
mu2 <- 0
mu3 <- 0
mu4 <- 0
sigma <- 1
alpha <- 0.05
K <- 4
```

```{r, echo=FALSE}
sim_results1 <- bonferroni_simulation(n_sims, n, mu1, sigma, K, alpha)
sim_results2 <- bonferroni_simulation(n_sims, n, mu2, sigma, K, alpha)
sim_results3 <- bonferroni_simulation(n_sims, n, mu3, sigma, K, alpha)
sim_results4 <- bonferroni_simulation(n_sims, n, mu4, sigma, K, alpha)
combined_results <- data.frame(covers1 = sim_results1$covers, covers2 = sim_results2$covers, covers3 = sim_results3$covers, covers4 = sim_results4$covers)
combined_results$covers_all <- combined_results$covers1 & combined_results$covers2 & combined_results$covers3 & combined_results$covers4
```

```{r, echo = FALSE}
cat("Combined coverage probability: ", mean(combined_results$covers_all), "\n")
cat("Group 1 coverage: ", mean(sim_results1$covers), "\n")
cat("Group 2 coverage: ", mean(sim_results2$covers), "\n")
cat("Group 3 coverage: ", mean(sim_results3$covers), "\n")
cat("Group 4 coverage: ", mean(sim_results4$covers))
```
