---
title: "Augusta University: STAT 7630"
subtitle: "Applied Linear Models"
author: "Dustin Pluta"
date: "2024 JAN 09"
output:
  xaringan::moon_reader:
    css: [metropolis-fonts, "my-theme.css"]
    lib_dir: libs
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature:
      highlightStyle: haddock
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
```


# Two-way ANOVA

### Luteinizing Hormone Study

_Rice 12.5.23_

> For a study of the release of luteinizing hormone (LH), male and female rats kept in constant light were compared to male and female rats in a regime of 14 h of light and 10 h of darkness. Various dosages of luteinizing releasing factor (LRF) were given: control (saline), 10, 50, 250, and 1250 ng. Levels of LH (in nanograms per milliliter of serum) were measured in blood samples at a later time.  


---

# Two-way ANOVA

### Luteinizing Hormone Study

_Rice 12.5.23_

- LH plays an important role in regulation of estrogen and testosterone, and so may exhibit different effects in male and female rats.

- We wish to analyze the data to determine the effects of light regime and LRF on release of LH.

- Specifically, we want to know:

  + Considering male rats' data separately, do mean LH levels depend on light regime?
  
  + Do mean LH levels depend on LRF dosage?
  
  + Is there an interaction between LRF dosage and light regime?
  
---

# Two-way ANOVA

```{r, echo = TRUE}
dat_wide <- read_csv("/home/dustin/Dropbox/AU/Teaching/STAT7630/Lectures/Lectures_source/dat/datasets/Chapter 12/LHmale.csv")
dat_wide
```

---

# Two-way ANOVA

```{r}
dat <- dat_wide %>% gather(value = "LH", key = "Light", -Dose)
dat
```

---

# Two-way ANOVA

```{r}
dat_summary <- dat %>% group_by(Dose, Light) %>% summarize(LH_mean = mean(LH), LH_sd = sd(LH) / n(), upr = LH_mean + 2 * LH_sd, lwr = LH_mean - 2 * LH_sd)
dat_summary
```

---

# Two-way ANOVA

```{r}
ggplot(dat) + 
  geom_boxplot(aes(x = factor(Dose), y = LH, fill = factor(Dose))) + 
  ggtitle("Marginal Mean LH across Dose")
```

---

# Two-way ANOVA

```{r}
ggplot(dat) + 
  geom_boxplot(aes(x = Light, y = LH, fill = Light)) + 
  ggtitle("Marginal Mean LH across Light Type")
```

---

# Two-way ANOVA

```{r}
ggplot(dat) + 
  geom_boxplot(aes(x = factor(Dose), y = LH, fill = Light), alpha = 0.5) +
  ggtitle("LH Mean by Light-Dose Combination") +
  xlab("Dose")
```


---

# Two-way ANOVA

```{r}
ggplot(dat_summary) + 
  geom_segment(aes(x = factor(Dose), xend = factor(Dose), y = lwr, yend = upr, color = Light), size = 2, alpha = 0.75) +
  geom_line(aes(x = factor(Dose), y = LH_mean, group = Light, linetype = Light, color = Light), size = 2) + 
  geom_point(aes(x = factor(Dose), y = LH_mean, group = Light, fill = Light), size = 2.5, pch = 23) +
  ggtitle("Interaction Plot for LH Data") + 
  xlab("Dose")
```

---

# Two-way ANOVA

```{r}
fit_Dose <- lm(LH ~ factor(Dose), data = dat)
anova(fit_Dose)
```

---

# Two-way ANOVA


```{r}
fit_Light <- lm(LH ~ Light, data = dat)
anova(fit_Light)
```

---

# Two-way ANOVA

```{r}
fit_no_int <- lm(LH ~ factor(Dose) + Light, data = dat)
anova(fit_no_int)
```

---

# Two-way ANOVA

```{r}
fit_int <- lm(LH ~ factor(Dose) + Light + Light:factor(Dose), data = dat)
anova(fit_int)
```

---

```{r}
plot(fit_int$fitted, fit_int$residuals, ylab = "Residuals", xlab = "Fitted")
```

---

# Two-way ANOVA

```{r}
qqnorm(fit_int$residuals)
qqline(fit_int$residuals)
```

---

# Two-way ANOVA

```{r}
dat$LH_log <- log(dat$LH)
```

---

# Two-way ANOVA

```{r}
dat_summary <- dat %>% group_by(Dose, Light) %>% summarize(LH_mean = mean(LH_log), LH_sd = sd(LH_log) / n(), upr = LH_mean + 2 * LH_sd, lwr = LH_mean - 2 * LH_sd)
ggplot(dat_summary) + 
  geom_segment(aes(x = factor(Dose), xend = factor(Dose), y = lwr, yend = upr, color = Light), size = 2, alpha = 0.75) +
  geom_line(aes(x = factor(Dose), y = LH_mean, group = Light, linetype = Light, color = Light), size = 2) + 
  geom_point(aes(x = factor(Dose), y = LH_mean, group = Light, fill = Light), size = 2.5, pch = 23) +
  ggtitle("Interaction Plot for LH Data") + 
  xlab("Dose")
```

---

# Two-way ANOVA

```{r}
fit_int_log <- lm(LH_log ~ factor(Dose) + Light + Light:factor(Dose), data = dat)
anova(fit_int_log)
```

---

# Two-way ANOVA

```{r}
plot(fit_int_log$fitted, fit_int_log$residuals, ylab = "Residuals", xlab = "Fitted")
```

---

# Two-way ANOVA

```{r}
qqnorm(fit_int_log$residuals)
qqline(fit_int_log$residuals)
```

